---
title: "Digital Humanities Text Analysis"
format: html
editor: visual
---

## Introduction

This Quarto notebook introduces basic text analysis workflow using five texts from my Polymath project. I used Project Gutenberg to gather the texts. The goal is not to learn every part of R syntax. Instead, we will follow what happens to a text as it moves through a text analysis workflow. Below are the four texts used:

- Johann Wolfgang von Goethe, Theory of Colours

- Alexander von Humboldt, Cosmos

- Mary Somerville, Connexion of the Physical Sciences

- Ibn Khaldun, The Muqaddimah

- Al-Biruni, Alberuni's India Vol. 1

  To analyze, I followed these steps in order to get the results.\
  \

  1)  Load Packages
  2)  Load text files
  3)  Organized the files
  4)  Tokenize text
  5)  Remove stop words
  6)  Count word frequencies
  7)  Visualize recurring themes
  8)  Interpret the data
      1)  When interpreting the data, the question I asked is what patterns can analysis help us notice, and what do we still need to return to the original text to understand?

## 1) Load Packages

R Packages add tools for particular kinds of tasks.\
For this project, we are using packages for working with data, reading text, cleaning text, text analysis, and visualization. Remember to load the tools before trying to execute the code to perform the analysis. tidyverse works with data and visualization and tidytext is specifically for text analysis

```{r}
library(tidyverse)
library(tidytext)
```

## 2) Load Texts

The texts have already been downloaded from Project Gutenberg and stored as plain-text files in the data_raw folder. What this code is doing is asking R to read each file and store its contents.\
File organization is important in understanding or finding the correct files. I had certain versions of the Humbolt and Muqaddimah, so I labeled them with ' vol1' and 'FINAL'.

The substr command shows me the first 500 characters of the text.

```{r}
goethe <- read_file("data_raw/goethe_colours.txt")
substr(goethe, 1, 500)

humboldt <- read_file("data_raw/humboldt_cosmos_vol1.txt")

somerville <- read_file("data_raw/somerville_connexion.txt")

muqaddimah <- read_file("data_raw/muqaddimah_FINAL.txt")

al_biruni <- read_file("data_raw/al_biruni_india_vol1.txt")

```

## 3) Building the Corpus (Putting texts together)

Combine Texts

To compare the authors' works, we created a table that connects each author with their text. After we run this tibble() code, the texts can be analyzed together while keeping track of which words belong to which author.

```{r}
texts <- tibble(
  author = c(
    "Goethe",
    "Humboldt",
    "Somerville",
    "Ibn Khaldun",
    "Al-Biruni"
  ),
  text = c(
    goethe,
    humboldt,
    somerville,
    muqaddimah,
    al_biruni
  )
)
texts
```

## 4) Tokenize Texts

Tokenization breaks text into smaller units called 'tokens'. Here, each token is a word -\
For example: 'Nature is a living whole' becomes:\
nature\
is\
a\
living\
whole

Breaking the text apart allows R to count and compare individual words. The words(language) are being turned into data that can be counted.

```{r}
words <- texts %>%
  unnest_tokens(word, text)
head(words, 10)
```

This is the point where we're really turning the books into analyzable data. unnest_tokens() breaks the text into individual words. And head() simply allows us to look at the first ten rows of the result.

## 5) Examine and Remove Stop Words

Words such as "the", "and", "of", "to" and "in" occur frequently but often add little value for analysis. For this analysis, we remove common stop words so other vocabulary becomes easier to see.\
\
This is a researcher decision based on the choice of the texts being studied and the researcher. Removing stop words is a methodological choice. Some common words may matter, depending on the context.

```{r}
data("stop_words")

head(stop_words, 10)

```

stop_words is an existing list of words commonly excluded from some text analyses. Looking at the list first reminds us that removing these words is a methodological decision

```{r}
words_clean <- words %>%
  anti_join(stop_words, by = "word")
head(words_clean, 10)

```

## 6) Normalize variations of certain words

```{r}
words_clean <- words_clean %>%
  mutate(
    word = case_when(
      word %in% c(
        "colour",
        "colours",
        "coloured",
        "color",
        "colors",
        "colored"
      ) ~ "colour",
      TRUE ~ word
    )
  )
```

R doesn't understand that colour, colours, coloured, and the American spelling 'color' are related concepts. To R, they're simply different strings of characters. For my research question, I'm more interested in references to the concept of colour or color, rather than differences between these grammatical forms or spellings. So, I'm telling R to treat these variants as one term: colour. This is called **'normalization'.**

```{r}
words_clean %>%
  filter(word == "colour") %>%
  count(author, word)
```

Remove project-specific noise After inspecting the results, I found tokens such as i.e and 1 that are not meaningful for this analysis. Standard stop-word lists do not necessarily identify every term that may need to be removed from a particular corpus.

```{r}
custom_stop_words <- c("i.e", "1")

words_clean <- words_clean %>%
  filter(!word %in% custom_stop_words)
```

then, I will double check to make sure:

```{r}
words_clean %>%
  filter(word %in% custom_stop_words)
```

it should say '0' rows, or something similar.

## 7) Count Word Frequencies

Now that the texts have been tokenized, stop words have been removed, and selected word variations have been normalized, we can count how often each word appears for each author. First, we create a table of word counts

```{r}
word_counts <- words_clean %>%
  count(author, word, sort = TRUE)

head(word_counts, 10)
```

count() counts how often each word appears for each author. Next, we keep the ten most frequent words for each author.

```{r}
top_words <- word_counts %>%
  group_by(author) %>%
  slice_max(n, n = 10)

top_words

```

## 8) Visualize word frequencies

ggplot2 is a visualization package included with the tidyverse. It builds visualizations in layers. In this graph: - top_words is the data we are visualizing - word provides the word labels - n is the number of times each word occurs - geom_col() creates the bars - coord_flip() turns the bars horizontally - facet_wrap() creates a separate panel for each author The + signs add each layer to the visualization.

```{r}
ggplot(top_words,aes(
    x = reorder(word, n),
    y = n,
    fill = author
  )) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~author, scales = "free")
```

What words stand out? What seems different between authors? What surprises you? What would you want to investigate further?

## 9) Return to texts and interpret (with patterns observed)

Text analysis can identify patterns, however, a frequency count doesn't discuss what words mean in context.\
\
I created questions that I think I would ask or discuss with the texts to continue the inquiry:

1.  Are there shared vocabularies among these authors?
2.  Which authors appear more concerned with nature?
3.  Which authors appear more concerned with society or history?
4.  What are the limitations of word-frequency analysis?
5.  How might translations affect patterns?
6.  Are unexpected words coming from the author, the translator, the editor, or the digital edition?

For example, we saw words such as al, de, and footnote. Instead of immediately deleting those words, the next step would be to return to the source and investigate why they appear
