---
title: "Computational Text Analysis"
format: html
editor: visual
---

## Introduction

This Quarto notebook introduces basic computational text analysis using R. I used Project Gutenberg to gather the texts. I analyzed four polymath texts:

- Goethe, Theory of Colours

- Humboldt, Cosmos

- Mary Somerville, Connection of the Physical Sciences

- Ibn Khaldun, Muqaddimah\
  \
  To analyze, I followed these steps in order to get the results.\
  \

  1)  Load Packages\
  2)  Load text files\
  3)  Tokenize text\
  4)  Remove stop words\
  5)  Count word frequencies\
  6)  Visualize recurring themes

## 1) Load Packages

When you click the **Render** button a document will be generated that includes both content and the output of embedded code.

```{r}
library(tidyverse)
library(tidytext)
library(readr)
library(stringr)
```

You can add options to executable code like this

```{r}
#| echo: false
2 * 2
```

The `echo: false` option disables the printing of code (only output is displayed

## 2) Load Texts

The texts have already been downloaded from Project Gutenberg and stored in the data_raw folder.

```{r}
goethe <- read_file("data_raw/goethe_colours.txt")

humboldt <- read_file("data_raw/humboldt_cosmos_vol1.txt")

somerville <- read_file("data_raw/somerville_connexion.txt")

muqaddimah <- read_file("data_raw/muqaddimah_FINAL.txt")
```

## 3) Create a Data Frame

Combine Texts

To compare authors, we first place all texts into a single table.

```{r}
texts <- tibble(
  author = c(
    "Goethe",
    "Humboldt",
    "Somerville",
    "Ibn Khaldun"
  ),
  text = c(
    goethe,
    humboldt,
    somerville,
    muqaddimah
  )
)
```

## 4) Tokenize Texts

Tokenization breaks a text into individual words.

```{r}
words <- texts %>%
  unnest_tokens(word, text)
```

## 5) Remove Stop Words

Words such as "the", "and", and "of" occur frequently but often add little analytical value.

```{r}
data("stop_words")

words_clean <- words %>%
  anti_join(stop_words, by = "word")
```

## 6) Count Word Frequencies

This will count the most common words and how often they are used throughout the whole text.

```{r}
top_words <- words_clean %>%
  count(author, word, sort = TRUE) %>%
  group_by(author) %>%
  slice_head(n = 10)
```

## 7) Visualize recurring themes (images)

This is how we see all of the computational text analysis in action - with images of graphs and charts.

```{r}
ggplot(
  top_words,
  aes(
    x = reorder(word, n),
    y = n,
    fill = author
  )
) +
  geom_col(show.legend = FALSE) +
  coord_flip() +
  facet_wrap(~author, scales = "free")
```

## 8) Interpretive Questions over the actual research.

Discussion questions:\

1.  Which words dominate each text?
2.  Are there shared vocabularies among these authors?
3.  Which authors appear more concerned with nature?
4.  Which authors appear more concerned with society or history?
5.  What are the limitations of word-frequency analysis?
